---
title: "Statistical Inference Course Project (Part 2)"
author: "Annette Spithoven"
date: "3-11-2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## 1. Overview

The purpose of the this data analysis is to analyze the ToothGrowth data set. This dataset contains information guinea tooth growth by supplement and dose.
We will perform exploratory analysis and hypothesis testing to figure out is there a relationship between tooth growth and daily dosage of vitamin C through various supplements.

```{r}
## loading the data
data("ToothGrowth")

## make it an environment object
ToothGrowth <- ToothGrowth
```

## 2. Basic Summary of the Data
The basic summary of the data start with an examination of the data structure.
```{r }
str(ToothGrowth)
```
The supp variable exist of two levels, OJ and VC. We rename those so that they are clearer. In addition, we will take a look at the information in the dataframe.
```{r }
## rename the levels
levels(ToothGrowth$supp) <- c("Orange Juice", "Ascorbic Acid")

## provide summary
summary(ToothGrowth)
```

```{r }
ggplot(ToothGrowth, 
       aes(x=as.character(dose), 
           y=len, 
           fill = supp)) + 
  ## make boxplots
  geom_boxplot() +
    ## seperate by supplement
  facet_grid(.~supp) +
  ## change titles
  labs(title="Guinea pig Tooth Length by Dosage for each type of Supplement", 
    x="Dose (mg/day)",
    y="Tooth Length",
    fill = "Supplement")+
  theme_bw()
```

## Use confidence intervals & hypothesis tests to compare tooth growth by supplement and dose

For each dosage we can test whether the supplement affects the tooth length of guinea pigs. 

```{r }

```

